---
title: "qc check"
author: "MG"
output: html_document
date: "2024-01-18"
---

# Setup
```{r setup}
workdir = "/hpc/pmc_vanheesch/projects/mvangijzen/nbl_longread"
fastqdir = paste0(workdir, "/analysis/fastq_ingress_results")
analysisdir = paste0(workdir, "/analysis")
savedir = paste0(workdir, "/results/qc_stats")
```

# Libraries
```{r libraries}
library(magrittr)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(ggpubr)
```

# Load data
```{r assign file names}
A13_read_stats_tsv = paste0(fastqdir, "/A13/fastcat_stats/per-read-stats.tsv")
A14_read_stats_tsv = paste0(fastqdir, "/A14/fastcat_stats/per-read-stats.tsv")
A15_read_stats_tsv = paste0(fastqdir, "/A15/fastcat_stats/per-read-stats.tsv")
E02_read_stats_tsv = paste0(fastqdir, "/E02/fastcat_stats/per-read-stats.tsv")
E03_read_stats_tsv = paste0(fastqdir, "/E03/fastcat_stats/per-read-stats.tsv")
E05_read_stats_tsv = paste0(fastqdir, "/E05/fastcat_stats/per-read-stats.tsv")
```

```{r read files}
# read length
A13_read_length <- data.table::fread(file = A13_read_stats_tsv,
                                     select=c(4:5))
A14_read_length <- data.table::fread(file = A14_read_stats_tsv,
                                     select=c(4:5))
A15_read_length <- data.table::fread(file = A15_read_stats_tsv,
                                     select=c(4:5))
read_length_df <- rbind(A13_read_length,A14_read_length, A15_read_length)

# read length entinostat
E02_read_length <- data.table::fread(file = E02_read_stats_tsv,
                                     select=c(4:5))
E03_read_length <- data.table::fread(file = E03_read_stats_tsv,
                                     select=c(4:5))
E05_read_length <- data.table::fread(file = E05_read_stats_tsv,
                                     select=c(4:5))
E_read_length_df <- rbind(E02_read_length,E03_read_length, E05_read_length)

# read length combined
com_read_length_df <- rbind(A13_read_length,A14_read_length, A15_read_length,
                            E02_read_length,E03_read_length, E05_read_length)

# mean quality
A13_mean_quality <- data.table::fread(file = A13_read_stats_tsv,
                                     select=c(4,6))
A14_mean_quality <- data.table::fread(file = A14_read_stats_tsv,
                                     select=c(4,6))
A15_mean_quality <- data.table::fread(file = A15_read_stats_tsv,
                                     select=c(4,6))
mean_quality_df <- rbind(A13_mean_quality,A14_mean_quality, A15_mean_quality)

# mean quality entinostat
E02_mean_quality <- data.table::fread(file = E02_read_stats_tsv,
                                     select=c(4,6))
E03_mean_quality <- data.table::fread(file = E03_read_stats_tsv,
                                     select=c(4,6))
E05_mean_quality <- data.table::fread(file = E05_read_stats_tsv,
                                     select=c(4,6))
E_mean_quality_df <- rbind(E02_mean_quality,E03_mean_quality, E05_mean_quality)

# select both
A13_stats <- data.table::fread(file = A13_read_stats_tsv,
                                     select=c(4:6))
A14_stats <- data.table::fread(file = A14_read_stats_tsv,
                                     select=c(4:6))
A15_stats <- data.table::fread(file = A15_read_stats_tsv,
                                     select=c(4:6))

```

```{r load efficiently}
#create filelist
filelist <- list(A13_read_stats_tsv, A14_read_stats_tsv,
                 A15_read_stats_tsv, E02_read_stats_tsv,
                 E03_read_stats_tsv, E05_read_stats_tsv)
#create read length df
readlength_all <- data.frame()

for (f in filelist){
#create tmp df for individual sample
tmp <- data.table::fread(file = f, select=c(4:6))
#add tmp df to readlength_all df
readlength_all <- rbind(readlength_all, tmp)
rm(tmp)
}
```

```{r add condition}
readlength_all$condition <- ifelse(readlength_all$sample_name %in%
                                   c("A13", "A14","A15"),
                                   "control","entinostat")
```

# Remove outliers
## Based on interquartile range
```{r remove outliers example code}
dim(readlength_all)
## [1] 247990240         4

quartiles <- quantile(readlength_all$read_length, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(readlength_all$read_length)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
readlength_no_outlier <- subset(readlength_all, 
                                readlength_all$read_length > Lower & 
                                readlength_all$read_length < Upper)
 
dim(readlength_no_outlier)
## [1] 242467624         4
```

# Read Length
## A13 remove outliers
```{r remove outliers example code}
dim(A13_read_length)
## [1] 247990240         3

quartiles <- quantile(A13_read_length$read_length, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(A13_read_length$read_length)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
A13_read_length_rm_outliers <- subset(A13_read_length, 
                                A13_read_length$read_length > Lower & 
                                A13_read_length$read_length < Upper)
 
dim(A13_read_length_rm_outliers)
## [1] 242467624         3
```

## A13 read length distribution
```{r}
A13_readlength_mean <- mean(A13_read_length$read_length)

A13length_plot <- A13_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(A13_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 500000, 
             label = paste0("mean = ", format(round(A13_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 550000, 
             label = paste0("n = ", length(A13_read_length$sample_name)))

A13length_plot
```

## A13 save read length plot
```{r save plot}
ggsave(filename = "A13_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## A14 remove outliers
```{r remove outliers example code}
dim(A14_read_length)
## [1] 247990240         3

quartiles <- quantile(A14_read_length$read_length, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(A14_read_length$read_length)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
A14_read_length_rm_outliers <- subset(A14_read_length, 
                                A14_read_length$read_length > Lower & 
                                A14_read_length$read_length < Upper)
 
dim(A14_read_length_rm_outliers)
## [1] 242467624         3
```

## A14 read length distribution
```{r}
A14_readlength_mean <- mean(A14_read_length$read_length)

A14length_plot <- A14_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(A14_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 500000, 
             label = paste0("mean = ", format(round(A14_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 550000, 
             label = paste0("n = ", length(A14_read_length$sample_name)))

A14length_plot
```

## A14 save read length plot
```{r save plot}
ggsave(filename = "A14_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## A15 remove outliers
```{r remove outliers example code}
dim(A15_read_length)
## [1] 247990240         3

quartiles <- quantile(A15_read_length$read_length, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(A15_read_length$read_length)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
A15_read_length_rm_outliers <- subset(A15_read_length, 
                                A15_read_length$read_length > Lower & 
                                A15_read_length$read_length < Upper)
 
dim(A15_read_length_rm_outliers)
## [1] 242467624         3
```

## A15 read length distribution
```{r}
A15_readlength_mean <- mean(A15_read_length$read_length)

A15length_plot <- A15_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(A15_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 500000, 
             label = paste0("mean = ", format(round(A15_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 600000, 
             label = paste0("n = ", length(A15_read_length$sample_name)))

A15length_plot
```

## A15 save read length plot
```{r save plot}
ggsave(filename = "A15_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## Compared read length distribution
```{r}
readlength_plot <- read_length_df %>%
  ggplot(aes(sample_name, read_length)) +
    geom_boxplot(fill="darkcyan") +
    #coord_cartesian(ylim = c(0,6000)) +
    labs(x="Sample", y="Read length (bp)") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

readlength_plot
```

## E02 read length distribution
```{r}
E02_readlength_mean <- mean(E02_read_length$read_length)

E02length_plot <- E02_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(E02_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 400000, 
             label = paste0("mean = ", format(round(E02_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 425000, 
             label = paste0("n = ", length(E02_read_length$sample_name)))

E02length_plot
```

## E02 save read length plot
```{r save plot}
ggsave(filename = "E02_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## E03 read length distribution
```{r}
E03_readlength_mean <- mean(E03_read_length$read_length)

E03length_plot <- E03_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(E03_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 400000, 
             label = paste0("mean = ", format(round(E03_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 450000, 
             label = paste0("n = ", length(E03_read_length$sample_name)))

E03length_plot
```

## E03 save read length plot
```{r save plot}
ggsave(filename = "E03_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## E05 read length distribution
```{r}
E05_readlength_mean <- mean(E05_read_length$read_length)

E05length_plot <- E05_read_length %>%
  ggplot(aes(read_length)) +
    geom_histogram(binwidth = 20, fill = "darkcyan") +
    coord_cartesian(xlim = c(0, 7500)) +
    labs(x = "Read length (bp)", y = "Count") +
    theme_classic() +
    geom_vline(xintercept = mean(E05_read_length$read_length), 
               color = "black", linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 4500, y = 400000, 
             label = paste0("mean = ", format(round(E05_readlength_mean, 2), nsmall = 2))) +
    annotate("text", x = 4500, y = 460000, 
             label = paste0("n = ", length(E05_read_length$sample_name)))

E05length_plot
```

## E05 save read length plot
```{r save plot}
ggsave(filename = "E05_readlength.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## Entinostat compared read length distribution
```{r}
E_readlength_plot <- E_read_length_df %>%
  ggplot(aes(sample_name, read_length)) +
    geom_boxplot(fill="darkcyan") +
    coord_cartesian(ylim = c(0,6000)) +
    labs(x="Sample", y="Read length (bp)") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

E_readlength_plot
```

## Entinostat vs normal read length distribution
```{r}
all_readlength_plot <- readlength_no_outlier %>%
  ggplot(aes(condition, read_length)) +
    geom_boxplot(aes(colour = factor(condition))) +
    #coord_cartesian(ylim = c(0,6000)) +
    labs(y="Read length (bp)") +
    scale_colour_manual(values = c("control" = "darkcyan",
                                   "entinostat" = "darkorange2")) +
    theme_classic() +
  theme(legend.title = element_blank()) +
    scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          axis.title.x = element_blank())

all_readlength_plot
```

# Mean Quality
## A13 mean quality distribution
```{r}
A13_quality_mean <- mean(A13_mean_quality$mean_quality)

A13quality_plot <- A13_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.1,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count") +
    geom_vline(xintercept = A13_quality_mean, color = "black", 
               linetype = "dashed", linewidth = 0.6) +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 25, y = 400000, 
             label = paste0("mean = ", format(round(A13_quality_mean, 2), nsmall = 2))) +
    annotate("text", x = 25, y = 425000, 
             label = paste0("n = ", length(A13_mean_quality$sample_name)))

A13quality_plot
```

## A13 save mean quality plot
```{r save plot}
ggsave(filename = "A13_meanquality.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## A14 mean quality distribution
```{r}
A14_quality_mean <- mean(A14_mean_quality$mean_quality)

A14quality_plot <- A14_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.1,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count") +
    geom_vline(xintercept = A14_quality_mean, color = "black", 
               linetype = "dashed", linewidth = 0.6) +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 25, y = 400000, 
             label = paste0("mean = ", format(round(A14_quality_mean, 2), nsmall = 2))) +
    annotate("text", x = 25, y = 425000, 
             label = paste0("n = ", length(A14_mean_quality$sample_name)))

A14quality_plot
```

## A14 save mean quality plot
```{r save plot}
ggsave(filename = "A14_meanquality.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## A15 mean quality distribution
```{r}
A15_quality_mean <- mean(A15_mean_quality$mean_quality)

A15quality_plot <- A15_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.1,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count") +
    geom_vline(xintercept = A15_quality_mean, color = "black", 
               linetype = "dashed", linewidth = 0.6) +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 25, y = 400000, 
             label = paste0("mean = ", format(round(A15_quality_mean, 2), nsmall = 2))) +
    annotate("text", x = 25, y = 425000, 
             label = paste0("n = ", length(A15_mean_quality$sample_name)))

A15quality_plot
```

## A15 save mean quality plot
```{r save plot}
ggsave(filename = "A15_meanquality.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## Compared quality distribution
```{r}
meanquality_plot <- mean_quality_df %>%
  ggplot(aes(sample_name, mean_quality)) +
    geom_boxplot(fill="darkcyan") +
    coord_cartesian(ylim = c(7,22)) +
    labs(x="Sample", y="Mean quality") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

meanquality_plot
```

## Untreated mean quality boxplot
```{r save plot}
ggsave(filename = "untreated_meanquality_boxplot_zoom.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

## E02 mean quality distribution
```{r}
E02quality_plot <- E02_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.2,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count",
         title = "E02 mean quality distribution") +
    geom_vline(xintercept = mean(E02_mean_quality$mean_quality), 
                color="red", linetype="solid") +
    geom_vline(xintercept = median(E02_mean_quality$mean_quality), 
                color="black", linetype="dashed") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

E02quality_plot
```

## E03 mean quality distribution
```{r}
E03quality_plot <- E03_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.2,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count",
         title = "E03 mean quality distribution") +
    geom_vline(xintercept = mean(E03_mean_quality$mean_quality), 
                color="red", linetype="solid") +
    geom_vline(xintercept = median(E03_mean_quality$mean_quality), 
                color="black", linetype="dashed") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

E03quality_plot
```

## E05 mean quality distribution
```{r}
E05quality_plot <- E05_mean_quality %>%
  ggplot(aes(mean_quality)) +
    geom_histogram(binwidth=0.2,fill="darkcyan") +
    geom_density(alpha=0.6) +
    coord_cartesian(xlim = c(0,30)) +
    labs(x="Mean quality", y="Count",
         title = "E05 mean quality distribution") +
    geom_vline(xintercept = mean(E05_mean_quality$mean_quality), 
                color="red", linetype="solid") +
    geom_vline(xintercept = median(E05_mean_quality$mean_quality), 
                color="black", linetype="dashed") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

E05quality_plot
```

## Entinostat compared quality distribution
```{r}
E_meanquality_plot <- E_mean_quality_df %>%
  ggplot(aes(sample_name, mean_quality)) +
    geom_boxplot(fill="darkcyan") +
    coord_cartesian(ylim = c(0,30)) +
    labs(x="Sample", y="Mean quality") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    # scale_x_continuous(expand = c(0, 0)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

E_meanquality_plot
```

# Mean quality per read length category
minimum read length A13 = 11
maximum read length A14 = 117238
```{r divide read length in categories}
A13_stats$length_category <- cut(A13_stats$read_length, 
                                 breaks = c(-Inf, 500, 1000, 1500, 2000, Inf),
                                 labels = c("[0-500]", "[500-1000]", "[1000-1500]", "[1500-2000]", "[2000<]"))
```

```{r plot mean quality per category}
until500 <- subset(A13_stats, length_category=="[0-500]")
until500_median <- median(until500$mean_quality)

A13_quality_per_length <- A13_stats %>%
  ggplot(aes(length_category, mean_quality)) +
    geom_boxplot(fill="darkcyan") +
    coord_cartesian(ylim = c(9,20)) +
    labs(x="Length category", y="Mean quality") +
    geom_abline(slope=0,intercept = until500_median, linetype="dashed") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) 
    # scale_x_continuous(expand = c(0, 0)) +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1))

A13_quality_per_length
```

## Saving 
```{r save plot}
ggsave(filename = "A13_quality_per_length_category_zoom.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```

# Pychopper stats
## Load pychopper files
```{r load files}
A13_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/A13_pychopper.tsv")
A14_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/A14_pychopper.tsv")
A15_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/A15_pychopper.tsv")
E02_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/E02_pychopper.tsv")
E03_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/E03_pychopper.tsv")
E05_pychop_stats_tsv = paste0(workdir, "/analysis/pychopper_stats/E05_pychopper.tsv")
```

```{r load efficiently}
#create filelist
filelist <- list(A13_pychop_stats_tsv, A14_pychop_stats_tsv,
                 A15_pychop_stats_tsv, E02_pychop_stats_tsv,
                 E03_pychop_stats_tsv, E05_pychop_stats_tsv)
#create read length df
pychop_stats <- data.frame()

for (f in filelist){
#create tmp df for individual sample
tmp <- data.table::fread(file = f,
                         nrows=10)
#add tmp df to readlength_all df
pychop_stats <- rbind(pychop_stats, tmp)
rm(tmp)
}

# subset for classification
pychop_class <- subset(pychop_stats, pychop_stats$Category == "Classification")

pychop_class$Name <- factor(pychop_class$Name, levels=c("Unusable", 
                                                        "Rescue",
                                                        "Primers_found"))
```

## Plot stats
```{r plot}
plot_class <- ggplot(pychop_class, 
                aes(fill=sample_id, Name, Value)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x="Classification", y="Number of reads") +
  scale_fill_brewer(palette = "Dark2") +
  #scale_fill_manual(values = c("darkgray", "orange", "royalblue",
   #                            "red", "darkgreen", "black")) +
  theme_classic() +
  theme(legend.title = element_blank()) +
  scale_y_continuous(expand = c(0, 0))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_class
```

```{r plot}
plot_class2 <- ggplot(pychop_class, 
                aes(fill=Name, sample_id, Value)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x="Sample", y="Number of reads") +
  scale_fill_brewer(palette = "Dark2") +
  #geom_text(aes(label=Value), vjust=0, color="white", size=3.5)+
  theme_classic() +
  theme(legend.title = element_blank()) +
  scale_y_continuous(expand = c(0, 0))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_class2
```

```{r calculate percentage}
pychop_total <- subset(pychop_stats, pychop_stats$Name == "PassReads", 
                       select = c(3:4))

pychop_total <- pychop_total %>% 
  dplyr::rename(total = Value)

pychop_perc <- merge(pychop_class, pychop_total, by = "sample_id")

pychop_perc$percentage <- pychop_perc$Value / pychop_perc$total * 100
```

```{r plot percentage}
plot_classperc <- ggplot(pychop_perc, 
                aes(fill=Name, sample_id, percentage)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x="Sample", y="Percentage of reads") +
  scale_fill_brewer(palette = "Dark2") +
  geom_text(aes(label=round(percentage, digits = 1)),
            vjust=-0.2, color="white", size=3.5)+
  theme_classic() +
  theme(legend.title = element_blank()) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     expand = c(0, 0))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_classperc
```

```{r combine plots}
ggarrange(plot_class2, plot_classperc, 
          common.legend = TRUE, 
          legend = "bottom")
```
## Saving 
```{r save plot}
ggsave(filename = "pychop_classes_v2.pdf",
       device = "pdf",
       path = savedir,
       width = 8,
       height = 5
       )
```


# Basic Stats per Sample
## Load files
```{r assign file names}
A13_file_stats_tsv = paste0(fastqdir, "/A13/fastcat_stats/per-file-stats.tsv")
A14_file_stats_tsv = paste0(fastqdir, "/A14/fastcat_stats/per-file-stats.tsv")
A15_file_stats_tsv = paste0(fastqdir, "/A15/fastcat_stats/per-file-stats.tsv")
E02_file_stats_tsv = paste0(fastqdir, "/E02/fastcat_stats/per-file-stats.tsv")
E03_file_stats_tsv = paste0(fastqdir, "/E03/fastcat_stats/per-file-stats.tsv")
E05_file_stats_tsv = paste0(fastqdir, "/E05/fastcat_stats/per-file-stats.tsv")
```

```{r read files}
A13_file_stats <- data.table::fread(A13_file_stats_tsv, select = c(2:7))
A14_file_stats <- data.table::fread(A14_file_stats_tsv, select = c(2:7))
A15_file_stats <- data.table::fread(A15_file_stats_tsv, select = c(2:7))
E02_file_stats <- data.table::fread(E02_file_stats_tsv, select = c(2:7))
E03_file_stats <- data.table::fread(E03_file_stats_tsv, select = c(2:7))
E05_file_stats <- data.table::fread(E05_file_stats_tsv, select = c(2:7))

all_file_stats <- rbind(A13_file_stats, A14_file_stats, A15_file_stats,
                        E02_file_stats, E03_file_stats, E05_file_stats)

all_file_stats %>%
  rename(
    n_reads = n_seqs
  )
```

```{r create table}
#save table
qcres_dir = "/hpc/pmc_vanheesch/projects/mvangijzen/nbl_longread/results/qc_stats"
write.csv(all_file_stats, file = paste0(qcres_dir,"/per_sample_stats.csv"), sep = "")
```
